---
title: "MADPH_EMS"
author: "Charlotte Chase"
date: "7/15/2020"
output: html_document
---

```{r}
library(tidyverse)
library(devtools)
library(readxl)
library(viridis)
library(fastDummies)
library(scales)
```

#reading in EMS information

```{r}
ems <- read_excel(
  "MA_EMS_Opioid.xlsx",
  na = c("***", "(1-4)")
) 
```

#This is clearly a wide dataset, and we want to pivot to a long format to make it tidy.
#ncol() gives the number of columns in dataframe


# we supply a pattern, and grepl looks for that pattern in a given string. (ex...looking for "female" in "age.group")
#returns true if it finds female, false otherwise

#now we've created a column that says "male," "female", or "total", depending on the age.group column. 

#Next, get rid of extraneous info tagged onto the age group.--that's the second mutate.

#first, we're overriding age.group with whatever the "ifelse" statement evaluates.
#string detec (str_detect)-- you give it a string and a pattern, and it returns true if it finds that pattern in that string, or false otherwise.
```{r}
ems1 <- ems %>% 
  pivot_longer(cols = 2:ncol(ems), names_to = "age.group", values_to = "count") %>% 
  mutate( #checking if the word female is in the value "age.group"
    gender = ifelse( #making a new column, "gender". if it finds "female," then it makes the row female
      grepl("Female", age.group, fixed = TRUE), 
      "Female",
      ifelse( #if it isn't female, it checks to see if it's male. if it's male, then the row is called "male"
        grepl("Male", age.group, fixed = TRUE),
        "Male",
        "Total" #otherwise, it's called total
      )
    )
  ) %>% 
  mutate(
    age.group = ifelse(
      str_detect(age.group, "[0-9]*(\\+|\\-)([0-9]*|)"), #string detect...we're looking for a range with a dash, or a number with a +
      str_extract(age.group, "[0-9]*(\\+|\\-)([0-9]*|)"), #if we do find that pattern, use string extract
      "Total" #if it doesn't detect that, put the word "Total".
    )
  )

```
#look up R's regex documentation


#bar chart vs line graph??? think about it. 
#mess around with this. 


#fill = what categorical data should be broken up into different colors. Breaking up by "longitudinal" and by "one-off," because that's how we've done the counts.

#use "fill," because the color of the bars is being applied by the "fill" argument in the aesthetic

#categorical data is more useful for grouping the data in different ways...coloring it or faceting it, etc. (changing the shape of the groups, etc.) 
#categories here are...gender and age.group

#we have two categories of data
#better to facet by a smaller category
```{r, warning=FALSE, message=FALSE, error=FALSE}
ggplot(ems1)+
  aes(
    x = Year, 
    y = count,
    color = age.group
  )+
  labs(x = "Year", y = "EMS Calls for Suspected OD", title = "MA Opioid-Related EMS Incidents by year, 2013-2019")+
  geom_line()+
  guides(color = guide_legend(title = "Age Group"))+ #changing the title of legend
  scale_y_continuous(labels = comma, limits =c(0, 25000))+
  scale_x_continuous(limits = c(2013, 2019), breaks = seq(2013,2019,1))+ #seq(start, end, interval)
  theme(axis.title=element_text(size=14,face="bold"))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  theme(panel.spacing = unit(1.5, "lines"))+
  theme(axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)))+
  theme(axis.title.x = element_text(margin = margin(t = 20, r = 0, b = 0, l = 0)))+
  theme(plot.title = element_text(margin = margin(t = 0, r = 0, b = 20, l = 0)))+
  facet_wrap(~gender) #by gender
```
There are different cases where a column can be more useful than a row...


#can facet by gender as well....think about what works, and then we can make it fancy. 

grab a weird sentence...that has the time in military time and try to build a regex that will capture just the millitary time. use the string extract function for practice. 

#Rshiny time...enabling / disabling age groups.